---
title: "Comparing Measures of Causal Relevance"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 5
    code-fold: true
    include-in-header:
      text: |
        <script>
        window.MathJax = {
          tex: { macros: {
            llbracket: "[\\mkern-3.5mu[",
            rrbracket: "]\\mkern-3.5mu]"
          }}
        };
        </script>
execute:
  warning: false
  message: false
---

```{r imports}
#| code-summary: "Imports"

library(tidyverse)
library(afex)
library(glue)
library(brms)
library(loo)
library(knitr)
library(faintr)
library(cspplot)
library(patchwork)

# these options help Stan run faster
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")

# use the CSP-theme for plotting
theme_set(theme_csp())

# global color scheme from CSP
project_colors <- cspplot::list_colors() |> pull(hex)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = project_colors)
}
```


We are exploring differnet ways to formally define **Causal Relevance**

$$CS(C,S)$$
where $S$ is a sentence in context $C$.

# Causal Mutual Information Measures

Causal mutual information is defined as:
$$cMI (X,Y) = \sum_{x,y}q(x)p(y\mid do(x))\log_2\frac{p(y\mid do(x))}{p(y)}$$

Kinney and Lombrozo (2024) assume that all variables are intervened on, and as such the intervention prior is flat. The effect $Y$ is binary, e.g. blue wings/not blue wings. For the sake of generality let $Y=\{0,1\}$. We can then write:

$$cMI (X,Y) = \frac{1}{|X|} \left(\sum_{x}p(Y=1\mid do(x))\log_2\frac{p(Y=1\mid do(x))}{p(Y=1)} +\sum_{x}p(Y=0\mid do(x))\log_2\frac{p(Y=0\mid do(x))}{p(Y=0)} \right),$$

which corresponds to how `cMI` is calculated in the code below and closely mirrors the computation shown in Kinney and Lombrozo (2024).
```{r cMI}
#| code-fold: false
cMI <- function(probs, grouping) {
  grouped_probs <- sapply(split(probs, grouping), mean)

  probs1 <- grouped_probs # probabilities that, e.g. Bricofly will develop blue wings
  probs0 <- 1 - grouped_probs # probabilities that, e.g. Bricofly will **not** develop blue wings

  q_c <- 1 / length(probs1) #q(c) is simply a constant, since q is a uniform probability distribution

  p1 <- 1 / length(probs1) * sum(probs1) # P(E=1)
  p0 <- 1 / length(probs0) * sum(probs0) # P(E=0)

  q_c * (sum(probs1 * log2(probs1 / p1)) + sum(probs0 * log2(probs0 / p0)))
}
```

### KL's Compressed Causal Models

$$CS_{KL}(C,S) = cMI(X,Y)-cMI(X_s,Y)$$
$X_s$ may be a null compression, in which case the measure simplifies to $CS_{KL}(C,S) = CMI(X,Y)$.

```{r KL}
#| code-fold: false
KL_CS <- function(probs, uncompressed_grouping, compressed_grouping = NULL) {
  # special case: null compression
  if (is.null(compressed_grouping)) {
    return(cMI(probs, uncompressed_grouping))
  }

  cMI(probs, uncompressed_grouping) - cMI(probs, compressed_grouping)
}
```

### MF's Causal Compression for Communciation

$$CS_{MF}(C,S) = \sum_{i}w_i\:cMI(X_i,Y)$$
```{r MF}
#| code-fold: false
MF_CS <- function(probs, grouping_list, weights = c(1)) {
  sum(weights * vapply(grouping_list, \(g) cMI(probs, g), numeric(1)))
}
```

Sentence S semantically denotes a partition of subsets of a subset of variables. For example, **high** sentence from Experiment 1 below, assuming X is a four-way categorical variable, can have partitioning $$\llbracket S_{high}\rrbracket = \{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\},$$ referred to as `finest` below, or
$$\llbracket S_{high} \rrbracket = \{\{x_1\},\{x_2,x_3,x_4\}\},$$
referred to as `one_vs_rest` below. The same goes for **low** sentences in Experiment 1, which also could be represented by either of the two partitions: $\llbracket S_{low}\rrbracket = \{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\}, \llbracket S_{low} \rrbracket = \{\{x_2\},\{x_1,x_3,x_4\}\}$

The grouping arguments in `cMI`, `KL_CS`, and `MF_CS` allow us to define the desired granularity. For example, comparing `finest` and `one_vs_rest` in `KL_CS` in the $x=85$ situation gives us:

```{r granularity-comparison}
#| code-fold: false
tibble(
  granularity = c("finest", "one_vs_rest"),
  KL_CS = c(
    KL_CS(c(.85, .7, .01, .01), c(1, 2, 3, 4)) |> round(2),
    KL_CS(c(.85, .7, .01, .01), c(1, 2, 2, 2)) |> round(2)
  )
) |>
  kable()
```

Kinney and Lombrozo (2024) use **only *finest* representations** in their model (see, for example, that the result for finest above corresponds to the calculation in footnote 10 in KL (2024)), meaning that for their measure it holds that $\llbracket S_{high} \llbracket = \{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\} = \llbracket S_{low} \rrbracket.$

- When $\llbracket S_{high} \rrbracket = \{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\},$ we are putting $\{x_1\}$ up against $\{x_2\},\{x_3\},\{x_4\},$ in which case $CS(C,S_{high})$ equals:

```{r}
#| code-fold: false
# (We have four groups and no way to pick out x_1, which is in group 1)
KL_CS(c(.85, .7, .01, .01), c(1, 2, 3, 4)) |> round(2)
```

- When $\llbracket S_{low} \rrbracket = \{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\},$ we are putting $\{x_2\}$ up against $\{x_2\},\{x_3\},\{x_4\},$ in which case $CS(C,S_{low})$ equals:

```{r}
#| code-fold: false
# (We have four groups and no way to pick out x_2, which is in group 2)
KL_CS(c(.85, .7, .01, .01), c(1, 2, 3, 4)) |> round(2)
```

As we can see, $CS_{KL}$ **cannot differentiate** between $S_{high}$ and $S_{low}$ **when using the *finest* representation**.

However, when we switch from using the `finest` representation in $CS_{KL}$ to the `one_vs_rest` representation **the situation changes and we can differentiate between the two**!

- When $\llbracket S_{high} \rrbracket = \{\{x_1\},\{x_2,x_3,x_4\}\},$ we are putting $\{x_1\}$ up against $\{x_2,x_3,x_4\},$ in which case  $CS(C,S_{high})$ equals:

```{r}
#| code-fold: false
# (We have two groups, x_1 is in group 1)
KL_CS(c(.85, .7, .01, .01), c(1, 2, 2, 2)) |> round(2)
```

- When $\llbracket S_{low} \rrbracket = \{\{x_2\},\{x_1,x_3,x_4\}\},$ we are putting $\{x_2\}$ up against $\{x_1,x_3,x_4\},$ in which case  $CS(C,S_{low})$ equals:

```{r}
#| code-fold: false
# (We have two groups, x_2 is in group 2)
KL_CS(c(.85, .7, .01, .01), c(1, 2, 1, 1)) |> round(2)
```


$CS_{MF}$ combines the two representations and so for $S_{high}$, with a probabilistic weight $p$, we get:

$$CS(C,S_{high}) = p \: cMI(\{\{x_1\},\{x_2\},\{x_3\},\{x_4\}\}, Y) + (1-p) \: cMI(\{\{x_1\},\{x_2,x_3,x_4\}\}, Y).$$




# Comparison: Kinney and Lombrozo (2024)

## Experiment 1

Table below is an overview of the measures applied to High and Low in Experiment 1.

```{r overview-table}
probs_70 <- c(.7, .7, .01, .01)
probs_85 <- c(.85, .7, .01, .01)
probs_98 <- c(.98, .7, .01, .01)


finest_grouping <- c(1, 2, 3, 4)
one_vs_rest_grouping <- c(1, 2, 2, 2)
compressed_grouping <- c(1, 1, 2, 2)

p <- .1
MF_weights <- c(p, 1 - p)
# MF_weights <- c(1,1)
results <- tibble(
  x = c("70", "85", "98"),
  'KL_finest_CS' = c(
    KL_CS(probs_70, finest_grouping) |> round(2),
    KL_CS(probs_85, finest_grouping) |> round(2),
    KL_CS(probs_98, finest_grouping) |> round(2)
  ),
  'KL_one_vs_rest_CS' = c(
    KL_CS(probs_70, one_vs_rest_grouping) |> round(2),
    KL_CS(probs_85, one_vs_rest_grouping) |> round(2),
    KL_CS(probs_98, one_vs_rest_grouping) |> round(2)
  ),
  'KL_compressed_CS' = c(
    KL_CS(probs_70, compressed_grouping) |> round(2),
    KL_CS(probs_85, compressed_grouping) |> round(2),
    KL_CS(probs_98, compressed_grouping) |> round(2)
  ),
  'KL_finest_CS - KL_compressed_CS' = c(
    KL_CS(probs_70, finest_grouping, compressed_grouping) |> round(2),
    KL_CS(probs_85, finest_grouping, compressed_grouping) |> round(2),
    KL_CS(probs_98, finest_grouping, compressed_grouping) |> round(2)
  ),
  'KL_one_vs_rest_CS - KL_compressed_CS' = c(
    KL_CS(probs_70, one_vs_rest_grouping, compressed_grouping) |> round(2),
    KL_CS(probs_85, one_vs_rest_grouping, compressed_grouping) |> round(2),
    KL_CS(probs_98, one_vs_rest_grouping, compressed_grouping) |> round(2)
  ),
  'MF_CS (weights: one_vs_rest {toString(MF_weights[1])}, finest {toString(MF_weights[2])})' := c(
    MF_CS(probs_70, list(one_vs_rest_grouping, finest_grouping), MF_weights) |>
      round(2),
    MF_CS(probs_85, list(one_vs_rest_grouping, finest_grouping), MF_weights) |>
      round(2),
    MF_CS(probs_98, list(one_vs_rest_grouping, finest_grouping), MF_weights) |>
      round(2)
  ),
  'MF_compressed_CS (weights: one_vs_rest {toString(MF_weights[1])}, finest {toString(MF_weights[2])})' := c(
    MF_CS(probs_70, list(compressed_grouping)) |> round(2),
    MF_CS(probs_85, list(compressed_grouping)) |> round(2),
    MF_CS(probs_98, list(compressed_grouping)) |> round(2)
  )
)

results |>
  column_to_rownames("x") |>
  t() |>
  kable()
```


To compare how well these measures model participant data, we first wrangle the data from Experiment 1 and initialise `KLexp1` with columns:

-   `pid` — Participant ID,
-   `vignette` — `Bricofly` / `Chapagite` / `Drol`; between-subjects,
-   `compression_kind` — `Prop` or `Stab`; between-subjects,
-   `claim_kind` — `Comp` / `High` / `Low`. Three kinds of statements visible at once on one screen,
-   `x_val` — Value of `x` in the vignette. Three levels: 70, 85, 98 which correspond to different amounts of loss; Between-subjects,
-   `participant_response` — participant's rating 'how good it would be to include each of the following statements in a summary of this report'; scale from -3 (very bad) to 3 (very good).

We then add predictions given by the different measures on each trial, fit cumulative logistic regression models and compare model fit.

`KLexp1`:
```{r KL-exp1-wrangling}
#| code-summary: "Data Wrangling"

rawexp1 <- read_csv(
  "KinneyLombrozo2024/Data/study_1_data.csv",
  col_types = cols(.default = col_character())
)

# glue('Raw data: {nrow(rawexp1)} rows, {ncol(rawexp1)} columns.')

# Drop the two Qualtrics label rows (question text and ImportId).
KLexp1df <- rawexp1 |> slice(-(1:2))

# Keep only participants who passed the comprehension check for their
# vignette. (%in% treats NA as a non-match, mirroring the original logic.)
#
# NOTE — bug fix: the original notebook looped over `range(2, len(data))`
# AFTER dropping two rows, so the final two participants in the file were
# never checked and were retained regardless of their comprehension-check
# answers. Here every participant is checked. If those last two rows
# happened to fail the check, sample sizes will differ slightly from the
# original output.
KLexp1df <- KLexp1df |>
  filter(
    D_CompCheck %in%
      "Drol" |
      B_CompCheck %in% "Bricofly" |
      C_CompCheck %in% "Chapagite"
  )

# glue('After comprehension check: {nrow(KLexp1df)} rows, {ncol(KLexp1df)} columns.')

# Remove all data from participants who rated any of the poor ("Bad")
# causal claims positively (i.e., gave a rating of 0, 1, 2, or 3).
# This replaces the original notebook's 54 near-identical filtering loops.
# As in the original, missing values do not trigger exclusion.
bad_cols <- grep(
  "^[BCD]_(Prop|Stab)(Comp|High|Low)Bad(70|85|98)_1$",
  names(KLexp1df),
  value = TRUE
)
stopifnot(length(bad_cols) == 54)

KLexp1df <- KLexp1df |>
  filter(!if_any(all_of(bad_cols), \(x) x %in% c("0", "1", "2", "3"))) |>
  mutate(pid = row_number())

# glue('After bad claim filtering: {nrow(KLexp1df)} rows, {ncol(KLexp1df)} columns.')

exp_cols <- grep("^[BCD]_(Prop|Stab)", names(KLexp1df), value = TRUE)

KLexp1_long <- KLexp1df |>
  pivot_longer(
    all_of(exp_cols),
    names_to = c(
      "prefix",
      "compression_kind",
      "claim_kind",
      "claim_valence",
      "x_val"
    ),
    names_pattern = "^([BCD])_(Prop|Stab)(Comp|High|Low)(Good|Bad)(70|85|98)_1$",
    names_transform = list(x_val = as.integer, rating = as.integer),
    values_to = "rating",
    values_transform = list(rating = as.integer),
    values_drop_na = TRUE
  )

KLexp1 <- KLexp1_long |>
  filter(claim_valence == "Good") |>
  rename(
    vignette = Vignette,
    participant_response = rating
  ) |>
  mutate(
    compression_kind = factor(compression_kind, levels = c("Prop", "Stab")),
    claim_kind = factor(claim_kind, levels = c("Comp", "High", "Low")),
    pid = factor(pid)
  ) |>
  select(
    pid,
    vignette,
    compression_kind,
    claim_kind,
    x_val,
    participant_response
  )


high_vs_rest <- c(1, 2, 2, 2)
low_vs_rest <- c(2, 1, 2, 2)

KLexp1 <- KLexp1 |>
  mutate(
    KL_finest_CS = map2_dbl(x_val, claim_kind, \(x, k) {
      grouping <- switch(
        k,
        Comp = compressed_grouping,
        High = finest_grouping,
        finest_grouping
      )
      KL_CS(c(x / 100, .7, .01, .01), grouping)
    }),
    KL_one_vs_rest_CS = map2_dbl(x_val, claim_kind, \(x, k) {
      grouping <- switch(
        k,
        Comp = compressed_grouping,
        High = high_vs_rest,
        low_vs_rest
      )
      KL_CS(c(x / 100, .7, .01, .01), grouping)
    }),
    MF_CS = map2_dbl(x_val, claim_kind, \(x, k) {
      grouping_list <- switch(
        k,
        omp = list(compressed_grouping),
        High = list(high_vs_rest, finest_grouping),
        list(low_vs_rest, finest_grouping)
      )
      MF_CS(c(x / 100, .7, .01, .01), grouping_list, MF_weights)
    })
  )

glimpse(KLexp1)
```

We can now make a plot which corresponds to Fig. 2 in the paper:

```{r KL-exp1-fig2}
KLexp1 |>
  filter(claim_kind != "Low") |>
  ggplot(aes(x = factor(x_val), y = participant_response, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind",
    title = "Participant data (Fig. 2. from KL (2024))"
  ) +
  theme_minimal()
```

And plotting also `low`, we get:

```{r KL-exp1-fig2-with-low}
KLexp1 |>
  ggplot(aes(x = factor(x_val), y = participant_response, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind",
    title = "Participant data (Fig. 2. from KL (2024) w/ Low)"
  ) +
  theme_minimal()
```

And we can visualise output values from the various CS variants:
```{r outputs-plot}
#| fig-height: 10
KL_finest_plot <- KLexp1 |>
  ggplot(aes(x = factor(x_val), y = KL_finest_CS, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  # scale_y_continuous(breaks = 0:3) +
  # coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "CS",
    fill = "Causal claim kind",
    title = "Values from KL_finest_CS"
  ) +
  theme_minimal()

KL_one_vs_rest_plot <- KLexp1 |>
  ggplot(aes(x = factor(x_val), y = KL_one_vs_rest_CS, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  # scale_y_continuous(breaks = 0:3) +
  # coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "CS",
    fill = "Causal claim kind",
    title = "Values from KL_one_vs_rest_CS"
  ) +
  theme_minimal()


MF_plot <- KLexp1 |>
  ggplot(aes(x = factor(x_val), y = MF_CS, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  # scale_y_continuous(breaks = 0:3) +
  # coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "CS",
    fill = "Causal claim kind",
    title = paste0(
      "Values from MF_CS  (weights: one_vs_rest ",
      toString(MF_weights[1]),
      ", finest ",
      toString(MF_weights[2]),
      ")"
    )
  ) +
  theme_minimal()

(KL_finest_plot /
  KL_one_vs_rest_plot /
  MF_plot) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

### Fitting cumulative logistic regression

```{r fit-setup}
#| code-summary: "Setup"
KLexp1 <- KLexp1 |>
  mutate(
    participant_response = ordered(participant_response, levels = -3:3),
    x_val = factor(x_val)
  )
```

<details>
<summary>See model fits</summary>

**KL finest**

```{r fit-kl-finest}
fit_KL_finest <- brm(
  participant_response ~ KL_finest_CS + (1 | pid),
  data = KLexp1,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_KL_finest",
  file_refit = "on_change"
)
summary(fit_KL_finest)
```

**KL one-vs-rest**

```{r fit-kl-one-vs-rest}
fit_KL_one_vs_rest <- brm(
  participant_response ~ KL_one_vs_rest_CS + (1 | pid),
  data = KLexp1,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_KL_one_vs_rest",
  file_refit = "on_change"
)
summary(fit_KL_one_vs_rest)
```

**MF (combination)**

```{r fit-kl-both}
# without set p
fit_MF <- brm(
  participant_response ~ KL_finest_CS + KL_one_vs_rest_CS + (1 | pid),
  data = KLexp1,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_MF",
  file_refit = "on_change"
)
summary(fit_MF)
```

</details>

#### Model comparison (LOO-CV)

```{r loo-comparison}

loos <- list(
  KL_finest = loo(fit_KL_finest, moment_match = TRUE),
  KL_one_vs_rest = loo(fit_KL_one_vs_rest, moment_match = TRUE),
  MF = loo(fit_MF, moment_match = TRUE)
)

# for (nm in names(loos)) {
#   cat("\n---", nm, "---\n")
#   print(pareto_k_table(loos[[nm]]))
# }

# sapply(loos, function(l) sum(pareto_k_values(l) > 0.7))

loo_comp <- loo_compare(loos)
loo_comp
```

**Lambert's z-test**
```{r lambert}
#| code-fold: false
1 - pnorm(-loo_comp[2, 1], loo_comp[2, 2])
1 - pnorm(-loo_comp[3, 1], loo_comp[3, 2])
```

The model comparison suggests that we have a much better model fit using $CS_{MF}$ than $CS_{KL}$ with the finest representation.

### Fitting cumulative logistic regression - No-low test
As an experiment, we also fit all the previous models to a dataset that does not contain *Low* sentences to match how how Fig. 2. in the original paper is ommitting them in the plot.

```{r fit-no-low-setup}
#| code-summary: "Setup"
KLexp1_no_low <- KLexp1 |>
  filter(claim_kind != 'Low')
```

<details>
<summary>See model fits</summary>

**KL finest**
```{r fit-kl-finest-no-low}
fit_KL_finest_no_low <- brm(
  participant_response ~ KL_finest_CS + (1 | pid),
  data = KLexp1_no_low,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_KL_finest_no_low",
  file_refit = "on_change"
)
summary(fit_KL_finest_no_low)
```

**KL one-vs-rest**
```{r fit-kl-one-vs-rest-no-low}
fit_KL_one_vs_rest_no_low <- brm(
  participant_response ~ KL_one_vs_rest_CS + (1 | pid),
  data = KLexp1_no_low,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_KL_one_vs_rest_no_low",
  file_refit = "on_change"
)
summary(fit_KL_one_vs_rest_no_low)
```

**MF (combination)**
```{r fit-kl-both-no-low}
fit_MF_no_low <- brm(
  participant_response ~ KL_finest_CS + KL_one_vs_rest_CS + (1 | pid),
  data = KLexp1_no_low,
  family = cumulative(), save_pars = save_pars(all = TRUE),
  file = "models/fit_MF_no_low",
  file_refit = "on_change"
)
summary(fit_MF_no_low)
```

</details>

#### Model comparison (LOO-CV)

```{r loo-comparison-no-low}

loos_no_low <- list(
  KL_finest_no_low = loo(fit_KL_finest_no_low, moment_match = TRUE),
  KL_one_vs_rest_no_low = loo(fit_KL_one_vs_rest_no_low, moment_match = TRUE),
  MF_no_low = loo(fit_MF_no_low, moment_match = TRUE)
)

# for (nm in names(loos)) {
#   cat("\n---", nm, "---\n")
#   print(pareto_k_table(loos[[nm]]))
# }

# sapply(loos, function(l) sum(pareto_k_values(l) > 0.7))

loos_no_low <- loo_compare(loos_no_low)
loos_no_low
```

**Lambert's z-test**
```{r lambert-no-low}
#| code-fold: false
1 - pnorm(-loos_no_low[2, 1], loos_no_low[2, 2])
1 - pnorm(-loos_no_low[3, 1], loos_no_low[3, 2])
```


We can see that when we omit the datapoints for *Low*, like in Fig. 2., $CS_{KL}$ with the finest representation fits the data similarly well as $CS_{MF}$ based on the model comparison. Here, the $CS_{KL}$ with one_vs_rest representation is a significantly worse fit.


