---
title: "Comparing Measures of Causal Relevance"
author: "Erik Zeiner"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r imports}
#| code-summary: "Imports"

library(tidyverse)
library(afex)
library(glue)
library(brms)
library(knitr)
```

# Causal Mutual Information Measures

## Kinney and Lombrozo (2024)

$$CMI (\mathbf{C},E) = \sum_{\mathbf{c},e}q(\mathbf{c})p(e\mid do(\mathbf{c}))\log_2\frac{p(e\mid do(\mathbf{c}))}{p(e)}$$

Kinney and Lombrozo (2024) assumes that all variables are intervened on, and as such probability distribution $q$ becomes uniform. The effect $E$ is binary, e.g. blue wings/not blue wings, for the sake of generality let $E=\{0,1\}$. We can then write

$$CMI (\mathbf{C},E) = \frac{1}{|\mathbf{C}|} \left(\sum_{\mathbf{c}}p(E=1\mid do(\mathbf{c}))\log_2\frac{p(E=1\mid do(\mathbf{c}))}{p(E=1)} +\sum_{\mathbf{c}}p(E=0\mid do(\mathbf{c}))\log_2\frac{p(E=0\mid do(\mathbf{c}))}{p(E=0)} \right),$$

which corresponds to how `CMI_KL_orig` is computed in the code.

```{r CMI_KL_orig}
#| code-fold: false
CMI_KL_orig <- function(prob_vector) {
  # prob_vector: e.g. c(.85, .7, .01, .01)

  probs1 <- prob_vector # probabilities that, e.g. Bricofly will develop blue wings
  probs0 <- 1 - prob_vector # probabilities that, e.g. Bricofly will **not** develop blue wings

  q_c <- 1 / length(probs1) #q(c) is simply a constant, since q is a uniform probability distribution

  p1 <- 1 / length(probs1) * sum(probs1) # P(E=1)
  p0 <- 1 / length(probs0) * sum(probs0) # P(E=0)

  q_c * (sum(probs1 * log2(probs1 / p1)) + sum(probs0 * log2(probs0 / p0)))
}

compress <- function(prob_vector) {
  # compressing a 4x4 probability table to a 2x2 table

  c(mean(prob_vector[1:2]), mean(prob_vector[3:4]))
}

loss_KL_orig <- function(prob_vector) {
  CMI_KL_orig(prob_vector) - CMI_KL_orig(compress(prob_vector))
}
```

```{r CMI_KL_orig_predictions}
#| code-summary: "Predictions table"
tibble(
  x = c("70", "85", "98"),
  CMI_KL_orig_not_compressed = c(
    CMI_KL_orig(c(.7, .7, .01, .01)) |> round(2),
    CMI_KL_orig(c(.85, .7, .01, .01)) |> round(2),
    CMI_KL_orig(c(.98, .7, .01, .01)) |> round(2)
  ),
  CMI_KL_orig_compressed = c(
    CMI_KL_orig(compress(c(.7, .7, .01, .01))) |> round(2),
    CMI_KL_orig(compress(c(.85, .7, .01, .01))) |> round(2),
    CMI_KL_orig(compress(c(.98, .7, .01, .01))) |> round(2)
  ),
  KL_orig_loss = c(
    loss_KL_orig(c(.7, .7, .01, .01)) |> round(2),
    loss_KL_orig(c(.85, .7, .01, .01)) |> round(2),
    loss_KL_orig(c(.98, .7, .01, .01)) |> round(2)
  )
) |>
  kable()
```

## Our approach: triple-C
TODO

## NSM
TODO

## CESM
TODO

# Comparison: Kinney and Lombrozo (2024)

### Experiment 1
We wrangle the data from Experiment 1, and initialise `KLexp1_measures` with data from the experiment and predictions given by `KL_orig` measure.

Base structure of the KLexp1_measures dataset is as follows:

-   `pid` — Participant ID.
-   `vignette` — `Bricofly` / `Chapagite` / `Drol`; between-subjects
-   `compression_kind` — `Prop` or `Stab`; between-subjects
-   `claim_kind` — `Comp` / `High` / `Low`. Three kinds of statements visible at once on one screen.
-   `x_val` — Value of `x` in the vignette. Three levels: 70, 85, 98 which correspond to different amounts of loss. Between-subjects.
-   `participant_response` — participant's rating 'how good it would be to include each of the following statements in a summary of this report'; scale from -3 (very bad) to 3 (very good)

This is then accompanied by prediction values given by various tested measures.

```{r KL-exp1-wrangling}
#| code-summary: "Data Wrangling"

rawexp1 <- read_csv(
  "KinneyLombrozo2024/Data/study_1_data.csv",
  col_types = cols(.default = col_character())
)

# glue('Raw data: {nrow(rawexp1)} rows, {ncol(rawexp1)} columns.')

# Drop the two Qualtrics label rows (question text and ImportId).
KLexp1df <- rawexp1 |> slice(-(1:2))

# Keep only participants who passed the comprehension check for their
# vignette. (%in% treats NA as a non-match, mirroring the original logic.)
#
# NOTE — bug fix: the original notebook looped over `range(2, len(data))`
# AFTER dropping two rows, so the final two participants in the file were
# never checked and were retained regardless of their comprehension-check
# answers. Here every participant is checked. If those last two rows
# happened to fail the check, sample sizes will differ slightly from the
# original output.
KLexp1df <- KLexp1df |>
  filter(
    D_CompCheck %in%
      "Drol" |
      B_CompCheck %in% "Bricofly" |
      C_CompCheck %in% "Chapagite"
  )

# glue('After comprehension check: {nrow(KLexp1df)} rows, {ncol(KLexp1df)} columns.')

# Remove all data from participants who rated any of the poor ("Bad")
# causal claims positively (i.e., gave a rating of 0, 1, 2, or 3).
# This replaces the original notebook's 54 near-identical filtering loops.
# As in the original, missing values do not trigger exclusion.
bad_cols <- grep(
  "^[BCD]_(Prop|Stab)(Comp|High|Low)Bad(70|85|98)_1$",
  names(KLexp1df),
  value = TRUE
)
stopifnot(length(bad_cols) == 54)

KLexp1df <- KLexp1df |>
  filter(!if_any(all_of(bad_cols), \(x) x %in% c("0", "1", "2", "3"))) |>
  mutate(pid = row_number())

# glue('After bad claim filtering: {nrow(KLexp1df)} rows, {ncol(KLexp1df)} columns.')

exp_cols <- grep("^[BCD]_(Prop|Stab)", names(KLexp1df), value = TRUE)

KLexp1_long <- KLexp1df |>
  pivot_longer(
    all_of(exp_cols),
    names_to = c(
      "prefix",
      "compression_kind",
      "claim_kind",
      "claim_valence",
      "x_val"
    ),
    names_pattern = "^([BCD])_(Prop|Stab)(Comp|High|Low)(Good|Bad)(70|85|98)_1$",
    names_transform = list(x_val = as.integer, rating = as.integer),
    values_to = "rating",
    values_transform = list(rating = as.integer),
    values_drop_na = TRUE
  )

KLexp1 <- KLexp1_long |>
  filter(claim_valence == "Good") |>
  rename(
    vignette = Vignette,
    participant_response = rating
  ) |>
  mutate(
    compression_kind = factor(compression_kind, levels = c("Prop", "Stab")),
    claim_kind = factor(claim_kind, levels = c("Comp", "High", "Low")),
    CMI_KL_orig = map2_dbl(x_val, claim_kind, \(x, k) {
      if (k == "Comp") {
        CMI_KL_orig(compress(c(x / 100, .7, .01, .01)))
      } else if (k == "High") {
        CMI_KL_orig(c(x / 100, .7, .01, .01))
      } else {
        CMI_KL_orig(c(.7, .7, .01, .01))
      }
    })
  ) |>
  select(
    pid,
    vignette,
    compression_kind,
    claim_kind,
    x_val,
    participant_response,
    CMI_KL_orig
  )

glimpse(KLexp1)
```

We can now make a plot which corresponds to Fig. 2 in the paper:

```{r KL-exp1-fig2}
KLexp1 |>
  filter(claim_kind != "Low") |>
  ggplot(aes(x = factor(x_val), y = participant_response, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind"
  ) +
  theme_minimal()
```

Adding `low`, we get:

```{r KL-exp1-fig2-with-low}

ggplot(
  KLexp1,
  aes(x = factor(x_val), y = participant_response, fill = claim_kind)
) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind"
  ) +
  theme_minimal()
```



TODO:

- add other measures
- fit cumulative logistic regression
- compare model fit
