---
title: "Comparing Measures of Causal Relevance"
author: "Erik Zeiner"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r imports}
#| code-summary: "Imports"

library(tidyverse)
library(afex)
library(glue)
```

# Kinney and Lombrozo (2024)

### Experiment 1

```{r KLexp1-processing}
#| code-summary: "Load & process data"

rawexp1 <- read_csv(
  "KinneyLombrozo2024/Data/study_1_data.csv",
  col_types = cols(.default = col_character())
)

# glue('Raw data: {nrow(rawexp1)} rows, {ncol(rawexp1)} columns.')

# Drop the two Qualtrics label rows (question text and ImportId).
KLexp1_wide <- rawexp1 |> slice(-(1:2))

# Keep only participants who passed the comprehension check for their
# vignette. (%in% treats NA as a non-match, mirroring the original logic.)
#
# NOTE — bug fix: the original notebook looped over `range(2, len(data))`
# AFTER dropping two rows, so the final two participants in the file were
# never checked and were retained regardless of their comprehension-check
# answers. Here every participant is checked. If those last two rows
# happened to fail the check, sample sizes will differ slightly from the
# original output.
KLexp1_wide <- KLexp1_wide |>
  filter(
    D_CompCheck %in%
      "Drol" |
      B_CompCheck %in% "Bricofly" |
      C_CompCheck %in% "Chapagite"
  )

# glue('After comprehension check: {nrow(KLexp1_wide)} rows, {ncol(KLexp1_wide)} columns.')

# Remove all data from participants who rated any of the poor ("Bad")
# causal claims positively (i.e., gave a rating of 0, 1, 2, or 3).
# This replaces the original notebook's 54 near-identical filtering loops.
# As in the original, missing values do not trigger exclusion.
bad_cols <- grep(
  "^[BCD]_(Prop|Stab)(Comp|High|Low)Bad(70|85|98)_1$",
  names(KLexp1_wide),
  value = TRUE
)
stopifnot(length(bad_cols) == 54)

KLexp1_wide <- KLexp1_wide |>
  filter(!if_any(all_of(bad_cols), \(x) x %in% c("0", "1", "2", "3"))) |>
  mutate(pid = row_number())

# glue('After bad claim filtering: {nrow(KLexp1_wide)} rows, {ncol(KLexp1_wide)} columns.')
```

The following function `loss` corresponds to $\mathscr{L}$ in the equation from Kinney and Lombrozo (2024):
$$\mathscr{L}(\mathbf{C}, \widehat{\mathbf{C}},E) = CMI(\mathbf{C},E) - CMI(\widehat{\mathbf{C}},E)$$
```{r loss-function}
#| code-fold: show

kl <- function(p, q) sum(p * log2(p / q))
# Kullback–Leibler divergence in bits, equivalent to `scipy.stats.entropy(p, q, base = 2)` in the original code

loss <- function(x) {
  conditional_noncomp1 <- c(.01, .99)
  conditional_noncomp2 <- c(.01, .99)
  conditional_noncomp3 <- c(.70, .30)
  conditional_noncomp4 <- c(x, 1 - x)
  conditional_comp1 <- c(.01, .99)
  conditional_comp2 <- c((.7 + x) / 2, 1 - (.7 + x) / 2)

  m <- .01 * .5 + .7 * .25 + x * .25
  marginal <- c(m, 1 - m)

  cmi_noncomp <- .25 *
    kl(conditional_noncomp1, marginal) +
    .25 * kl(conditional_noncomp2, marginal) +
    .25 * kl(conditional_noncomp3, marginal) +
    .25 * kl(conditional_noncomp4, marginal)
  cmi_comp <- .50 *
    kl(conditional_comp1, marginal) +
    .50 * kl(conditional_comp2, marginal)

  cmi_noncomp - cmi_comp
}
```

We construct the final processed dataframe for Experiment 1, `KLexp1`, including the loss values for each condition and the difference scores used in the original analysis:

- `pid` — Participant ID.
- `vignette` — `Bricofly` / `Chapagite` / `Drol`; between-subjects
- `compression_kind` — `Prop` or `Stab`; between-subjects
- `claim_kind` — `Comp` / `High` / `Low`. Three kinds of statements visible at once on one screen.
- `x_val` — Value of `x` in the vignette. Three levels: 70, 85, 98 which correspond to different amounts of loss. Between-subjects.
- `loss_val` — Loss value computed from `x_val` using the `loss` function above.
- `participant_response` — participant's rating 'how good it would be to include each of the following statements in a summary of this report'; scale from -3 (very bad) to 3
(very good)
- `comp_eval`, `high_eval`, `low_eval` — Evaluations of each claim type.
- `comp_high_diff` — `comp_eval - high_eval`
- `comp_avg_diff` — `comp_eval - (0.5 * high_eval + 0.5 * low_eval)`
- `high_low_diff` — `high_eval - low_eval`
- `feedback` — Feedback from participants

In addition, we create a separate dataframe `KLexp1_measures` which contains the columns needed for the comparison of causal relevance measures, including the original KL's loss values as `KL_measure_orig`:

```{r KL-exp1-wrangling}
#| code-summary: "Data Wrangling"

KLexp1_wide <- KLexp1_wide |>
  mutate(
    vignette_prefix = case_match(
      Vignette,
      "Bricofly" ~ "B",
      "Chapagite" ~ "C",
      "Drol" ~ "D"
    ),
    loss_level = case_match(
      Group,
      c("Prop_No_Loss", "Stab_No_Loss") ~ "70",
      c("Prop_Loss_1", "Stab_Loss_1") ~ "85",
      c("Prop_Loss_2", "Stab_Loss_2") ~ "98"
    ),
    loss_val = case_match(
      Group,
      c("Prop_No_Loss", "Stab_No_Loss") ~ loss(.70),
      c("Prop_Loss_1", "Stab_Loss_1") ~ loss(.85),
      c("Prop_Loss_2", "Stab_Loss_2") ~ loss(.98)
    ),
    mode_label = if_else(str_starts(Group, "Prop"), "Prop", "Stab")
  )

# Pull each participant's rating of a given claim from the column that
# matches their vignette / mode / loss condition.
pick_eval <- function(df, claim) {
  cols <- paste0(
    df$vignette_prefix,
    "_",
    df$mode_label,
    claim,
    "Good",
    df$loss_level,
    "_1"
  )
  m <- as.matrix(df[, unique(cols)])
  as.numeric(m[cbind(seq_len(nrow(df)), match(cols, colnames(m)))])
}

KLexp1_wide <- KLexp1_wide |>
  mutate(
    comp_eval = pick_eval(KLexp1_wide, "Comp"),
    high_eval = pick_eval(KLexp1_wide, "High"),
    low_eval = pick_eval(KLexp1_wide, "Low"),
    comp_high_diff = comp_eval - high_eval,
    comp_avg_diff = comp_eval - (0.5 * high_eval + 0.5 * low_eval),
    high_low_diff = high_eval - low_eval
  )

exp_cols <- grep("^[BCD]_(Prop|Stab)", names(KLexp1_wide), value = TRUE)

KLexp1_long <- KLexp1_wide |>
  pivot_longer(
    all_of(exp_cols),
    names_to = c(
      "prefix",
      "compression_kind",
      "claim_kind",
      "claim_valence",
      "x_val"
    ),
    names_pattern = "^([BCD])_(Prop|Stab)(Comp|High|Low)(Good|Bad)(70|85|98)_1$",
    names_transform = list(x_val = as.integer, rating = as.integer),
    values_to = "rating",
    values_transform = list(rating = as.integer),
    values_drop_na = TRUE
  )

KLexp1 <- KLexp1_long |>
  mutate(
    compression_kind = factor(compression_kind, levels = c("Prop", "Stab")),
    claim_kind = factor(claim_kind, levels = c("Comp", "High", "Low"))
  ) |>
  rename(
    vignette = Vignette,
    feedback = Feedback,
    participant_response = rating
  ) |>
  filter(claim_valence == "Good") |>
  select(
    pid,

    vignette,
    compression_kind,
    claim_kind,
    x_val,
    loss_val,

    participant_response,

    comp_eval,
    high_eval,
    low_eval,
    comp_high_diff,
    comp_avg_diff,
    high_low_diff,

    feedback
  )

KLexp1_measures <- KLexp1 |>
    select(pid, vignette, compression_kind, claim_kind, x_val, participant_response,loss_val) |>
    rename(
        KL_measure_orig = loss_val
    )
head(KLexp1_measures)
```

We can now make a plot which corresponds to Fig. 2 in the paper:
```{r KL-exp1-fig2}
KLexp1 |>
  filter(claim_kind != "Low") |>
  ggplot(aes(x = factor(x_val), y = participant_response, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind"
  ) +
  theme_minimal()
```

Adding `low` to Fig. 2:
```{r KL-exp1-fig2-with-low}

ggplot(KLexp1, aes(x = factor(x_val), y = participant_response, fill = claim_kind)) +
  stat_summary(
    fun = mean,
    geom = "col",
    width = 0.6,
    position = position_dodge(width = 0.7)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    position = position_dodge(width = 0.7)
  ) +
  scale_y_continuous(breaks = 0:3) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(
    x = "x (corresponds to loss level)",
    y = "Mean Evaluation of Causal Claim",
    fill = "Causal claim kind"
  ) +
  theme_minimal()
```






